---
title: "Using `glmBfp`: Cox models with test-based Bayes Factors"
output: rmarkdown::html_vignette
bibliography: examples.bib
vignette: >
  %\VignetteIndexEntry{Using `glmBfp`: Cox models with test-based Bayes Factors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(glmBfp)
```

The `glmBfp` package implements a new approach to model fitting and variable selection which are described in two articles, @held2015approximate and @held2016objective.
This vignette shows how to use the simplified interface for fitting Cox models with `glmBfp`.
We use the SMART data described in \citet{Steyerberg2009} and available at [https://www.clinicalpredictionmodels.org/]. The package includes a processed copy of this data with the missing values imputed and variables transformed as in the example code on that site

```{r}
data(SMARTfull)
```


The workhorse function of the package is `glmBayesMfp()` which does all the model fitting and returns a large list of possible models, which then need to be evaluated and selected from. To make things easier, `coxTBF()` is a simplified formula based interface to `glmBayesMfp()` that fits and chooses Cox models.
   
We first need to define the formula, using `Surv(time,event)`. The function needs to know which variables must be included in the model and which are "uncertain" should tested for inclusion. These are wrapped in the formula with `uc()`.

```{r formulas, cache=TRUE}
f1 <- Surv(TEVENT, EVENT) ~ AGE.TRANS + SEX + uc(SMOKING) +
  uc(ALCOHOL) + uc(BMI) + uc(SYSTH) + uc(HDL) + uc(DIABETES) +
  uc(HISTCAR2)
```
Using this we can fit a model. We choose `type="MAP"` to select the model with maximum posterior probability. Other possibilities are the median probability model (`MPM`) and Bayesian model averages (`BMA`).

```{r, message=FALSE, warning=FALSE, cache=TRUE}
f1_MAP <- coxTBF(f1,
  data = SMARTfull, type = "MAP",
  useOpenMP = FALSE, chainlength = 200
)
```

The resulting object contains the formula of selected model and the coefficient estimates.
```{r, cache = TRUE}
f1_MAP$formula
f1_MAP$coefs
```
Also included is the survivor function, so we can predict survival probabilities at specified times.

```{r, cache=TRUE}
predict(f1_MAP, times = c(100, 1000, 2000, 3000), newdata = SMARTfull[1:3, ])
```


Other parameters given to `coxTBF` are passed through to `glmBayesMfp`. This can be used to specify new $g$-priors and change MCMC options. We can also save the models found in the search for later investigation with `keepModelList=TRUE`. This time we select the MPM (median probability model). If this model is not one of the models found in the Monte Carlo search, then it is constructed and returned.

```{r, message=FALSE,  cache=TRUE, warning=FALSE}
# Hyper g/n.obs
prior <- InvGammaGPrior(a = 1 / 2, b = sum(SMARTfull$EVENT) / 2)
f1_MPM <- coxTBF(f1,
  data = SMARTfull, type = "MPM", useOpenMP = FALSE,
  chainlength = 500, nModels = 50, keepModelList = TRUE,
  priorSpecs = list(gPrior = prior, modelPrior = "sparse")
)
```

```{r, cache=TRUE}
f1_MPM$formula
f1_MPM$coefs
```




## References
